---
title: "RATA Analyses"
output:
  pdf_document:
    df_print: kable
    toc: yes
  html_notebook:
    theme: readable
    toc: yes
    # css: styles.css   #uncomment to use style sheet
    out.width: '100%'
  html_document:
    toc: yes
    # css: styles.css   #uncomment to use style sheet
    out.width: '100%'
---

# Summary
The following is an investigation of the historic use of alternate performance specifications (APS) for Relative Accuracy Test Audits (RATA) required under 40 CFR part 75, a statistical calculation of alternate performance specifications, a determination of revised alternate performance specifications, and an evaluation of the effect of the proposed alternate performance specifications on the overall pass/fail rate and frequency (annual/semi-annual) qualification of the historic RATA results.

The following RATA tests will be included in this analysis:

- SO2 and NOX concentration (ppm)
- NOX rate (lb/mmBtu)
- CO2 and O2 (percent)
- Moisture (%H2O)


The following provided data sets include data for the years 2014-2018:

- SO2, NOX concentration RATA (SO2RATA.csv and NOXRATA.csv)
- NOX Rate RATA (NOXRRATA.csv)
- CO2 and O2 RATA (CO2RATA.csv and O2RATA.csv)
- Moisture RATA (H2OMRATA.csv and H2ORATA.csv)

**To run this R Notebook, include the above files in the same directory as the R Notebook.**  
R requirements:  
R version 3.6.0  


The following packages were used in this analysis:  
*ggpubr_0.2.1, magrittr_1.5. kableExtra_1.1.0, DescTools_0.99.28, tinytex_0.14. knitr_1.23, data.table_1.12.2, forcats_0.4.0, stringr_1.4.0, dplyr_0.8.1, purr_0.3.2, readr_1.3.1, tidyr_0.8.3, tibble_2.1.3, ggplot2_3.2.0, tidyverse_1.2.1*    


```{r warning=FALSE, include=FALSE}
# load necessary packages
packages <- c("tidyverse", "data.table", "tidyr", "knitr", "tinytex", "DescTools", "kableExtra", "ggpubr")
missing.packages <- setdiff(packages,installed.packages()[,"Package"])
if(length(missing.packages)) install.packages(missing.packages)
for(pkg in packages) {
do.call("library",list(pkg))
}
```

```{r include=FALSE}
# set notebook options
options(scipen = 999)         # disable scientific notation
options(tibble.width = Inf)   # output width of tibbles
```

```{r }
# variables used in the analysis, can be changed as desired
ra_4qtr <- 7.5                # relative accuracy for annual reporting  
ra_2qtr <- 10                 # relative accuracy for semi-annual reporting
so2nox_4qtr_diff <- 12        # SO2 & NOx ppm APS value for annual reporting
so2nox_2qtr_diff <- 15        # SO2 & NOx ppm APS value for semi-annual reporting
noxr_4qtr_diff <- 0.015       # NOx rate lb/mmBtu APS value for annual reporting
noxr_2qtr_diff <- 0.200       # NOx rate lb/mmBtu APS value for semi-annual reporting
co2o2_4qtr_diff <- 0.7        # CO2 or O2 % APS value for annual reporting
co2o2_2qtr_diff <- 1.0        # CO2 or O2 % APS value for semi-annual reporting
h2o_4qtr_diff <- 1.0          # moisture (H2O) % APS value for annual reporting
h2o_2qtr_diff <- 1.5          # moisture (H2O) % APS value for semi-annual reporting
annual_aps <- .75             # percentage of proposed semi-annual APS which will determine the annual APS
```


```{r include=FALSE}
# The round function in R rounds to the nearest even number if the decimal value is .5, so we may not get correct values. 
# The following round function will be used instead.
# inputs:
# - x = value to be rounded
# - n = how many places to round
round2 <- function(x, n) (trunc((abs(x) * 10 ^ n) + 0.5) / 10 ^ n) * sign(x)
```

```{r include=FALSE}
# create function to read in files with appropriate error and warning messages
readfile <- function(file, ...) {
  tryCatch(read_csv(file, ...),
    error = function(e) message("File not found. Make sure file resides in directory with Rmd file.", file),
    warning = function(w) message("File not found. Make sure file resides in directory with Rmd file.", file)
  )
}   
```

```{r include=FALSE}
so2rata <- readfile("rawdata/SO2RATA.csv")
noxrata <- readfile("rawdata/NOXRATA.csv")
noxrrata <- readfile("rawdata/NOXRRATA.csv")
co2rata <- readfile("rawdata/CO2RATA.csv")
o2rata <- readfile("rawdata/O2RATA.csv")
h2omrata <- readfile("rawdata/H2OMRATA.csv")
h2orata <- readfile("rawdata/H2ORATA.csv")
```

```{r}
# function to check for missing values in Relative.Accuracy and Mean.Diff columns
checkmissingvals <- function(df) {
  df %>% summarise(missing_ra = sum(is.na(Relative.Accuracy)),
                   missing_mid = sum(is.na(Mean.Diff)))
}
```

```{r}
# check for missing values in so2rata
checkmissingvals(so2rata)
```

```{r}
# check for missing values in noxrata
checkmissingvals(noxrata)
```

```{r}
# check for missing values in noxrrata
# check for missing values in noxrata
checkmissingvals(noxrrata)
```

```{r}
# check for missing values in co2rata
checkmissingvals(co2rata)
```

```{r}
# check for missing values in o2rata
checkmissingvals(o2rata)
```

```{r}
# check for missing values in h2orata
checkmissingvals(h2orata)
```



# Analysis 
The historical data (2014-2018) will be used to determine unit status for meeting main performance requirements and alternate performance specification (APS) criteria. A statistical analysis will be performed on the units that pass to determine proposed annual and semi-annual APS. The initial data will then be used to test the proposed annual and semi-annual APS.


## Filter and combine files to prepare for analysis ##  
### SO2 and NOx (ppm) ###
Combine SO2 and NOX concentration RATA data (SO2RATA.csv and NOXRATA.csv).  
Only include RATA with a reference mean equal to or less than 250 ppm.
```{r }
# combine SO2 and NOX concentration RATA data
so2nox <- rbind(so2rata, noxrata)

# filter RATAs
so2nox_filtered <- so2nox %>% 
  filter(Mean.RATA.Reference <= 250)
```


### NOx Rate ###
From NOXRRATA.csv, only include RATAs with a reference mean equal to or less than 0.200 lb/mmBtu.
```{r include=FALSE}
# filter RATAs
noxrrata_filtered <- noxrrata %>% 
  filter(Mean.RATA.Reference <= 0.200)
```


### CO2 and O2 (%) ###
Combine CO2 and O2 RATA data (CO2RATA.csv and O2RATA.csv).
```{r include=FALSE}
# combine CO2 and O2 files and filter RATAs
co2o2 <- rbind(co2rata, o2rata)
```

### Moisture RATA (%H2O) ###
Combine H2O and H2OM data (H2ORATA.csv and H2OMRATA.csv).
```{r include=FALSE}
# combine H2O and H2OM files and filter RATAs
h2omh2o <- rbind(h2omrata, h2orata)
```


**-------------------------------------------------- FUNCTIONS --------------------------------------------------**
```{r include=FALSE}
theme_custom <- function (base_size = 10, base_family = "sans") {
  half_line <- base_size/2
  theme_light(base_size = base_size, base_family = base_family) %+replace%           # basic theme to start
  theme(
    line = element_line(colour = "black", 
                        size = 0.5, 
                        linetype = 1, 
                        lineend = "butt"), 
    rect = element_rect(fill = "white",                                              # background
                        colour = "black", 
                        size = 0.5, 
                        linetype = 1),
    text = element_text(family = base_family,                                        # text
                        face = "plain",
                        colour = "black", 
                        size = base_size,
                        lineheight = 0.9,  
                        hjust = 0.5,
                        vjust = 0.5, 
                        angle = 0, 
                        margin = margin(), 
                        debug = FALSE), 
    
    axis.line = element_blank(),                                                     # chart border
    axis.text = element_text(size = rel(0.8), colour = "grey20"),                    # axis text
    axis.text.x = element_text(margin = margin(t = 0.8*half_line/2), vjust = 1),     # axis x label text
    axis.text.y = element_text(margin = margin(r = 0.8*half_line/2), hjust = 1),     # axis y label text
    axis.ticks = element_line(colour = "grey20"),                                    # axis tick marks
    axis.ticks.length = unit(half_line/2, "pt"),                                     # axis tick mark length
    axis.title.x = element_text(size = rel(1),  face = "bold",                       # styles the x axis title
                                margin = margin(t = 10, r = 0, b = 0, l = 0)),
    axis.title.y = element_text(size = rel(1),  face = "bold", angle = 90,           # styles the y axis title
                                margin = margin(t = 0, r = 15, b = 0, l = 0)),
    
    legend.background = element_rect(fill = "grey92", colour = NA ),                 # legend background and border
    legend.spacing.x = unit(0.2, 'cm'),                                              # category spacing
    legend.key = element_rect(fill = "grey95", colour = "grey85"),                   # category background and border
    legend.key.size = unit(1.2, "lines"),                                            # category item size
    legend.key.height = NULL,                                                        # category item height
    legend.key.width = NULL,                                                         # category item width
    legend.text = element_text(size = rel(0.8)),                                     # legend category item text size
    legend.text.align = NULL,
    legend.title = element_text(hjust = 0), 
    legend.title.align = .5,                                                         # alignment of legend title (0-1 is left-right)
    legend.position = "bottom",                                                      # sets the location of the legend relative to the grid
    legend.direction = NULL,
    legend.justification = "center",                                                 # aligns the legend
    legend.box = NULL, 
    
    panel.background = element_rect(fill = "grey92", colour = NA),                   # sets the chart panel background
    panel.border = element_blank(),                                                  # sets the chart panel border
    panel.grid.major = element_line(colour = "white"),                               # sets the chart grid line
    panel.grid.minor = element_line(colour = "white", size = 0.25), 
    #panel.margin = unit(half_line, "pt"), 
    panel.margin.y = NULL, 
    panel.margin.x = NULL,
    panel.ontop = FALSE, 
    
    strip.background = element_rect(fill = "grey85", colour = NA),
    strip.text = element_text(colour = "grey10", size = rel(0.8)),
    strip.text.x = element_text(margin = margin(t = half_line, b = half_line)), 
    strip.text.y = element_text(angle = -90, 
                                margin = margin(l = half_line, r = half_line)),
    strip.switch.pad.grid = unit(0.1, "cm"),
    strip.switch.pad.wrap = unit(0.1, "cm"), 
    
    plot.background = element_rect(colour = "white"),                               # sets the figure background color
    plot.title = element_text(size = rel(1.2), face = "bold",                       # styles the chart title
                              hjust = 0.5, vjust = 2.5,
                              margin = margin(b = half_line * 1.2)),
    plot.margin = margin(half_line, half_line, half_line, half_line),

    complete = TRUE)
}
```

```{r }
# function to determine historical test results
# case_when is evaluated in order of statements
# required function inputs:  
# - df = filtered dataframe  
# - ra_4qtr - relative accuracy for annual evaluation
# - ra_2qtr - relative accuracy for semi-annual evaluation
# - diff_4qtr - mean difference between CEM and reference value for annual evaluation
# - diff_2qtr - mean difference between CEM and reference value for semi-annual evaluation
# output: a new dataframe with a column for testresults

testresults <- function(df, ra_4qtr, ra_2qtr, diff_4qtr, diff_2qtr) {
  df %>%
  mutate(testresult = case_when(
    Relative.Accuracy <= ra_4qtr ~ "PASS_4QTRS",
    Relative.Accuracy > ra_4qtr & between(Mean.Diff, -diff_4qtr, diff_4qtr) ~ "PASSAPS_4QTRS",
    Relative.Accuracy > ra_4qtr & Relative.Accuracy <= ra_2qtr ~ "PASS_2QTRS",
    Relative.Accuracy > ra_2qtr & between(Mean.Diff, -diff_2qtr, diff_2qtr) ~ "PASSAPS_2QTRS",
    TRUE ~ "FAILED"
    )
  )  
  
}
```



```{r }
# function to get statistics for RATA mean difference
# required function inputs:  
# - x = vector of values 
# output: statistical values for supplied vector
getstats <- function(x) { 
  stats_table <- data.table(
    mean = mean(x),
    se = MeanSE(x),
    median = median(x),
    mode = paste(Mode(x), collapse=", "),
    sd = sd(x),
    variance = var(x),
    kurtosis = Kurt(x),
    skewness = Skew(x),
    range = Range(x),
    min = min(x),
    max = max(x),
    sum = sum(x),
    count = length(x),
    sd1_low = mean(x) - sd(x),
    sd1_high = mean(x) + sd(x),
    sd1_pct = length(which(x >= (mean(x) - sd(x)) & 
                             x <= (mean(x) + sd(x)))) / length(x) * 100,
    sd2_low = mean(x) - (2*sd(x)),
    sd2_high = mean(x) + (2*sd(x)),
    sd2_pct = length(which(x >= (mean(x) - (2*sd(x))) & 
                             x <= (mean(x) + (2*sd(x))))) / length(x) * 100,
    sd3_low = mean(x) - (3*sd(x)),
    sd3_high = mean(x) + (3*sd(x)),
    sd3_pct = length(which(x >= (mean(x) - (3*sd(x))) & 
                             x <= (mean(x) + (3*sd(x))))) / length(x) * 100
    ) 
  
}
```

```{r}
# function to display statistics 
# required function inputs:  
# - df = dataframe of values for which you want statistics  
# output: statistical values for supplied variable displayed in a table
displaystats <- function(df) {
  df %>% mutate(sd1_range = str_c("(",sd1_low,", ",sd1_high,")"),
                 sd2_range = str_c("(",sd2_low,", ",sd2_high,")"),
                 sd3_range = str_c("(",sd3_low,", ",sd3_high,")")) %>%
  select(mean, "standard error" = se, median, mode, "standard deviation" = sd, variance, 
         kurtosis, skewness, range, "minimum" = min, "maximum" = max, sum, 
         "number of RATAs" = count, "1 standard deviation" = sd1_range, 
         "% RATAs in 1 standard deviation" = sd1_pct, "2 standard deviations" = sd2_range, 
         "% RATAs in 2 standard deviations" = sd2_pct, "3 standard deviations" = sd3_range, 
         "% of RATAs in 3 standard deviations" = sd3_pct) %>%
  gather(variable, value) %>% 
  kable() %>% 
  kable_styling(position="left")
}
```

```{r }
# function to add a column in each dataframe of passed observations to indicate the RATAs 
# standard deviation. This will be used to color the histograms.
# required function inputs:  
# - df = dataframe
# - dfstats = dataframe of statistical values for df (from getstats function)
# output: dataframe modified to include standard deviations column
addsd <- function(df, dfstats) {
df %>% mutate(sd = factor(case_when(
    Mean.Diff >= dfstats$sd1_low & Mean.Diff <= dfstats$sd1_high ~ "SD1",
    Mean.Diff >= dfstats$sd2_low & Mean.Diff <= dfstats$sd2_high ~ "SD2",
    Mean.Diff >= dfstats$sd3_low & Mean.Diff <= dfstats$sd3_high ~ "SD3",
    TRUE ~ "OTHER"
    ), levels = c("SD1","SD2","SD3","OTHER"))) %>% 
    arrange(sd) -> df
}
```

```{r }
# function to plot Mean and Standard Deviations of Mean Difference
# required function inputs:  
# - df = dataframe
# - dfstats = dataframe of statistical values for df (from getstats function)
# - bins = # of bins for histogram
# output: dataframe modified to include standard deviations column
plotstats <- function(df, dfstats, bins) {
  # arrange values in sd column in order for appearance in legend
gplot <- ggplot(df, aes(x=Mean.Diff, fill = sd)) + 
  geom_histogram(bins=bins, color="#3b3b3b", size = 0.5) + 
  geom_vline(aes(xintercept=mean(Mean.Diff)), color="blue", linetype="solid", size=1)  +
  xlab("Mean Difference") + 
  ylab("Frequency") + 
    theme_set(theme_custom()) +
  scale_fill_manual(values=c("#7AC36A","#5A9BD4","#FAA75B","#F15A60"), 
                    limits=c("SD1","SD2","SD3","OTHER")) +
  labs(fill="Standard Deviations")

}
```

```{r}
# create function to plot distribution with boxplot and histogram
# required function inputs:  
# - df = filtered dataframe 
# - var = variable used in plot
# - bins = # of bins for histogram
# - title - title for the plot
# output: a box plot and histogram plot of supplied variable
plotdist <- function(df,var,bins,title) {
  var <- enquo(var)   # convert unquoted var to a quoted one
  p1 <- ggplot(df, aes(x = "", y = !! var)) +    # !! required to unquote var
    stat_boxplot(geom = "errorbar", width = 0.4) + 
    geom_boxplot(outlier.colour = "#F15A60") +
    coord_flip() + 
    stat_summary(fun.y = mean, colour = "#9E67AB", geom = "point", shape = 18, size = 3) + 
    labs(x = "", y = "") + 
    theme_set(theme_custom()) + 
    ggtitle(title)
  
 
  p2 <- ggplot(df, aes(x=Mean.Diff, fill = sd)) + 
  geom_histogram(bins=bins, color="#3b3b3b", size = 0.5) + 
  geom_vline(aes(xintercept=mean(Mean.Diff)), color="blue", linetype="solid", size=1)  +
  xlab("Mean Difference") + 
  ylab("Frequency") + 
    theme_set(theme_custom()) +
  scale_fill_manual(values=c("#7AC36A","#5A9BD4","#FAA75B","#F15A60"), 
                    limits=c("SD1","SD2","SD3","OTHER")) +
  labs(fill="Standard Deviations")
  

   ggarrange(p1,p2,heights = c(.75, 2), align = "v", ncol = 1, nrow = 2)
}
```



**---------------------------------------------------------------------------------------------------------------**

## Historical Test Results ## 
Using the above filtered dataframes, we will determine the frequency of units Passing (meeting Relative Accuracy requirements), Passing with Alternative Performance Specification (meeting mean difference between CEMS and reference values), and Failing as well as determining the reporting frequency status (annually or semi-annually).

The following criteria will be utilized to determine if a RATA passes and the frequency:

* For SO2 and NOx ppm RATAs    
    + if Relative Accuracy is <= `r ra_4qtr `%, then RATA = PASS with a frequency of 4QT
    + if Relative Accuracy is > `r ra_4qtr `% and if the difference between the CEMS and reference value is within +/- `r so2nox_4qtr_diff ` ppm, then RATA = PASSAPS with a frequency of 4QTR
    + if Relative Accuracy is > `r ra_4qtr `% and <= `r ra_2qtr `%, then RATA = PASS with a frequency of 2QTR
    + if Relative Accuracy is > `r ra_2qtr `% and if the difference between the CEMS and reference value is within +/- `r so2nox_2qtr_diff ` ppm, then RATA = PASSAPS with a frequency of 2QTR  

* For NOx rate RATAs
    + if Relative Accuracy is <= `r ra_4qtr `%, then RATA = PASS with a frequency of 4QTR
    + if Relative Accuracy is > `r ra_4qtr `% and if the difference between the CEMS and reference value is within +/- `r noxr_4qtr_diff ` lb/mmBtu, then RATA = PASSAPS  with a frequency of 4QTR
    + if Relative Accuracy is > `r ra_4qtr `% and <= `r ra_2qtr `%, then RATA = PASS with a frequency of 2QTR
    + if Relative Accuracy is > `r ra_2qtr `% and if the difference between the CEMS and reference value is within +/- `r noxr_2qtr_diff ` lb/mmBtu, then RATA = PASSAPS with a frequency of 2QTR  

* For CO2 or O2 % RATAs
    + if Relative Accuracy is <= `r ra_4qtr `%, then RATA = PASS with a frequency of 4QTR
    + if Relative Accuracy is > `r ra_4qtr `% and if the difference between the CEMS and reference value is within +/- `r co2o2_4qtr_diff `%, then RATA = PASSAPS with a frequency of 4QTR
    + if Relative Accuracy is > `r ra_4qtr `% and <= `r ra_2qtr `%, then RATA = PASS with a frequency of 2QTR
    + if Relative Accuracy is > `r ra_2qtr `% and if the difference between the CEMS and reference value is within +/- `r co2o2_2qtr_diff `%, then RATA = PASSAPS with a frequency of 2QTR  

* For moisture RATAs
    + if Relative Accuracy is <= `r ra_4qtr `%, then RATA = PASS with a frequency of 4QTR
    + if Relative Accuracy is > `r ra_4qtr `% and if the difference between the CEMS and reference value is within +/- `r h2o_4qtr_diff `%, then RATA = PASSAPS with a frequency of 4QTR
    + if Relative Accuracy is > `r ra_4qtr `% and <= `r ra_2qtr `%, then RATA = PASS with a frequency of 2QTR
    + if Relative Accuracy is > `r ra_2qtr `% and if the difference between the CEMS and reference value is within +/- `r h2o_4qtr_diff `%, then RATA = PASSAPS with a frequency of 2QTR  


```{r }
# for each monitoring system, call the testresults function, supplying the filtered dataframe
# and the variables for relative accuracy and mean difference values
# function call: testresults(df,ra_4qtr,ra_2qtr,diff_4qtr,diff_2qtr)
so2nox_results <- testresults(so2nox_filtered, ra_4qtr, ra_2qtr, so2nox_4qtr_diff, so2nox_2qtr_diff)
noxr_results <- testresults(noxrrata_filtered, ra_4qtr, ra_2qtr, noxr_4qtr_diff, noxr_2qtr_diff)
co2o2_results <- testresults(co2o2, ra_4qtr, ra_2qtr, co2o2_4qtr_diff, co2o2_2qtr_diff)
h2o_results <- testresults(h2omh2o, ra_4qtr, ra_2qtr, h2o_4qtr_diff, h2o_2qtr_diff)
```


## Statistical Evaluation ## 
**Evaluation of the current Alternative Performance Specifications.**   

For each of the RATA tests:  

* exclude all observations that failed their tests  
* plot the frequency distribution of the mean difference  
* perform a basic statistical evaluation of the RATA mean difference listing the mean, standard error, median, mode, standard deviation, sample variance, kurtosis, skewness, range, minimum, maximum, sum, count, largest (x), smallest (X), confidence level, (X) percentage  
* plot the mean, standard deviation (1 sigma, 2 sigma, 3 sigma)    


### SO2 and NOx (ppm) ###
**Mean Difference for RATAs that passed**
```{r}
# keep only those observations that passed
so2nox_passed <- so2nox_results %>% filter(testresult !="FAILED") 
```


**Statistics for SO2 & NOx Mean Difference of RATAs that passed**
```{r warning=FALSE}
# statistical evaluation of all that passed
# function call: getstats(x)
so2nox_stats <- getstats(so2nox_passed$Mean.Diff) 
displaystats(so2nox_stats)
```

```{r }
# call addsd function to add column for standard deviation values for each RATA
# function call: addsd(df, dfstats)
so2nox_passed <- addsd(so2nox_passed, so2nox_stats)
```

```{r out.width='100%'}
# function call: plotdist(df, var, bins, title)
plotdist(so2nox_passed, Mean.Diff, 25, "SO2 & NOx ppm RATAs Mean Difference")
```


### NOx Rate ###
**Mean Difference for RATAs that passed**
```{r}
# keep only those observations that passed
noxr_passed <- noxr_results %>% filter(testresult !="FAILED")
```


**Statistics for NOx Rate Mean Difference of RATAs that passed**
```{r warning=FALSE}
# get statistics for NOx rate RATAs
# function call: getstats(x)
noxr_stats <- getstats(noxr_passed$Mean.Diff) 
displaystats(noxr_stats)
```

```{r }
# call addsd function to add column for standard deviation values for each RATA
# function call: addsd(df, dfstats)
noxr_passed <- addsd(noxr_passed, noxr_stats)
```

```{r out.width='100%'}
# plot frequency distribution of mean difference
# function call: plotdist(df, var, bins, title)
plotdist(noxr_passed, Mean.Diff, 25, "NOx Rate RATAs Mean Difference")
```

### CO2 and O2 (%) ###
**Mean Difference for RATAs that passed**
```{r}
# keep only those observations that passed
co2o2_passed <- co2o2_results %>% filter(testresult !="FAILED")
```


**Statistics for CO2 & O2 Mean Difference of RATAs that passed**
```{r warning=FALSE}
# function call: getstats(x)
co2o2_stats <- getstats(co2o2_passed$Mean.Diff) 
displaystats(co2o2_stats)
```

```{r }
# call addsd function to add column for standard deviation values for each RATA
# function call: addsd(df, dfstats)
co2o2_passed <- addsd(co2o2_passed, co2o2_stats)
```

```{r out.width='100%'}
# plot frequency distribution of mean difference
# function call: plotdist(df, var, bins, title)
plotdist(co2o2_passed, Mean.Diff, 25, "CO2 & O2 % RATAs Mean Difference")
```

### Moisture RATA (%H2O) ###
```{r}
# keep only those observations that passed
h2o_passed <- h2o_results %>% filter(testresult !="FAILED")
```


**Statistics for Moisture Mean Difference of RATAs that passed**
```{r warning=FALSE}
# function call: getstats(x)
h2o_stats <- getstats(h2o_passed$Mean.Diff) 
displaystats(h2o_stats)
```

```{r }
# call addsd function to add column for standard deviation values for each RATA
h2o_passed <- addsd(h2o_passed, h2o_stats)
```

```{r out.width='100%'}
# plot frequency distribution of mean difference
# function call: plotdist(df, var, bins, title)
plotdist(h2o_passed, Mean.Diff, 20, "Moisture (H2O) RATAs Mean Difference")
```

## Determination of proposed annual and semiannual APS ##
The proposed semi-annual APS will be the resulting greatest absolute value of 2 standard deviations from the mean rounded to the nearest whole value for SO2 and NOx ppm, to the nearest thousandths place for NOx rate, and nearest tenths place for CO2 or O2 percent, and moisture (%H2O).

```{r }
# get absolute value of 2 standard deviations from the mean for semi-annual frequency
so2nox_sa_aps <- round2(pmax(abs(so2nox_stats$sd2_low),abs(so2nox_stats$sd2_high)), 0)
noxr_sa_aps <- round2(pmax(abs(noxr_stats$sd2_low),abs(noxr_stats$sd2_high)),3)
co2o2_sa_aps <- round2(pmax(abs(co2o2_stats$sd2_low),abs(co2o2_stats$sd2_high)),1)
h2o_sa_aps <- round2(pmax(abs(h2o_stats$sd2_low),abs(h2o_stats$sd2_high)),1)

```

The proposed annual APS will be `r annual_aps ` of the proposed semi-annual APS rounded to the nearest whole value for SO2 and NOx ppm, to the nearest thousandths place for NOx rate, and nearest tenths place for CO2 or O2 percent, and moisture (% H2O).
```{r }
# get annaul APS values (25% of the proposed semi-annual APS value)
so2nox_a_aps <- round2(annual_aps*so2nox_sa_aps, 0) 
noxr_a_aps <- round2(annual_aps*noxr_sa_aps, 3)
co2o2_a_aps <- round2(annual_aps*co2o2_sa_aps, 1)
h2o_a_aps <- round2(annual_aps*h2o_sa_aps, 1)
```

```{r }
# create dataframe of test results
proposedaps <- data.frame("Annual" = c(so2nox_a_aps, noxr_a_aps, co2o2_a_aps, h2o_a_aps), 
                          "Semiannual" = c(so2nox_sa_aps, noxr_sa_aps, co2o2_sa_aps, h2o_sa_aps), 
                          row.names = c("SO2 NOx", "NOx Rate", "CO2 or O2", "Moisture"))

proposedaps %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("bordered","striped", "hover", "condensed", "responsive"), full_width = FALSE, position="left") %>% 
  row_spec(row=0, extra_css = "background-color: #2E8540; color: #fff;") %>%
  column_spec(column = 1, bold = TRUE) %>%
  add_header_above(c("Proposed APS" = 3), extra_css = "background-color: #0071BC; color: #fff;") #%
```

## Determination of pass and frequency rates with proposed APS ##  
The proposed alternate performance specifications will be evaluated on the historic RATA data to determine the overall pass/fail rate and frequency (annual/semi-annual) qualification.
```{r }
# create variables for the different proposed measurements
# variable values are proposed APS values calculated above
so2nox_prop_4qtr_diff <- so2nox_a_aps  
so2nox_prop_2qtr_diff <- so2nox_sa_aps  
noxr_prop_4qtr_diff <- noxr_a_aps  
noxr_prop_2qtr_diff <- noxr_sa_aps  
co2o2_prop_4qtr_diff <- co2o2_a_aps  
co2o2_prop_2qtr_diff <- co2o2_sa_aps  
h2o_prop_4qtr_diff <- h2o_a_aps  
h2o_prop_2qtr_diff <- h2o_sa_aps  
```


```{r }
## call function for each of the dataframes (excluding failed tests) with appropriate variables for relative accuracy
# and mean difference between CEMS and reference value
# function call: testresults(df,ra_4qtr,ra_2qtr,diff_4qtr,diff_2qtr)
so2nox_prop_results <- testresults(so2nox_passed, ra_4qtr, ra_2qtr, 
                                   so2nox_prop_4qtr_diff, so2nox_prop_2qtr_diff)
noxr_prop_results <- testresults(noxr_passed, ra_4qtr, ra_2qtr, 
                                 noxr_prop_4qtr_diff, noxr_prop_2qtr_diff)
co2o2_prop_results <- testresults(co2o2_passed, ra_4qtr, ra_2qtr, 
                                  co2o2_prop_4qtr_diff, co2o2_prop_2qtr_diff)
h2o_prop_results <- testresults(h2o_passed, ra_4qtr, ra_2qtr, 
                                h2o_prop_4qtr_diff, h2o_prop_2qtr_diff)
```

**Results**
```{r }
# Combine current and proposed dataframes
so2nox_comb_results <-add_column(so2nox_passed, propresults=so2nox_prop_results$testresult)
noxr_comb_results <-add_column(noxr_passed, propresults=noxr_prop_results$testresult)
co2o2_comb_results <-add_column(co2o2_passed, propresults=co2o2_prop_results$testresult)
h2o_comb_results <-add_column(h2o_passed, propresults=h2o_prop_results$testresult)
```

```{r }
# function to create a data table to display results
# required function inputs:  
# - df = dataframe
# output: dataframe of results of current and proposed APS
displayresults <- function(df) {
  data.table(
    numratas = df %>% nrow(),
    numpass4 = df %>% filter(testresult=="PASS_4QTRS") %>% nrow(),
    numpassaps4 = df %>% filter(testresult=="PASSAPS_4QTRS") %>% nrow(),
    numpass2 = df %>% filter(testresult=="PASS_2QTRS") %>% nrow(),
    numpassaps2 = df %>% filter(testresult=="PASSAPS_2QTRS") %>% nrow(),
    numproppass4 = df %>% filter(propresults=="PASS_4QTRS") %>% nrow(),
    numproppassaps4 = df %>% filter(propresults=="PASSAPS_4QTRS") %>% nrow(),
    numproppass2 = df %>% filter(propresults=="PASS_2QTRS") %>% nrow(),
    numproppassaps2 = df %>% filter(propresults=="PASSAPS_2QTRS") %>% nrow(),
    numpass = df %>% filter(!propresults=="FAILED") %>% nrow(),
    percpass = round2((df %>% filter(!propresults=="FAILED") %>% nrow())/(df %>% nrow())*100,2),
    numpasstofail = (df %>% filter(testresult !="FAILED") %>% filter(propresults=="FAILED") %>% nrow()),
    num4to2 = (df %>% filter(RATA.Frequency=="2QTRS" | RATA.Frequency=="4QTRS") %>% 
                 filter(testresult=="PASS_4QTRS" | testresult=="PASSAPS_4QTRS") %>%
                 filter(propresults=="PASS_2QTRS" | propresults=="PASSAPS_2QTRS") %>% nrow()),
    perc4to2 = round2(((df %>% filter(RATA.Frequency=="2QTRS" | RATA.Frequency=="4QTRS") %>%
                          filter(testresult=="PASS_4QTRS" | testresult=="PASSAPS_4QTRS") %>%
                          filter(propresults=="PASS_2QTRS" | propresults=="PASSAPS_2QTRS") %>% 
                                nrow()))/(df %>% nrow())*100,2)
    ) 
  
} 
```


```{r }
# results of all tests
# call displayresults function and then add a test column to denote the test for the data
so2nox_comparison <- so2nox_comb_results %>% 
  displayresults() %>% 
  mutate(test="so2nox") %>% 
  select(test, everything())
noxr_comparison <- noxr_comb_results %>% 
  displayresults() %>% 
  mutate(test="noxr") %>% 
  select(test, everything())
co2o2_comparison <- co2o2_comb_results %>% 
  displayresults() %>% 
  mutate(test="co2o2") %>% 
  select(test, everything())
h2o_comparison <- h2o_comb_results %>% 
  displayresults() %>% 
  mutate(test="h2o") %>% 
  select(test, everything())
```

```{r }
# combine above into one dataframe
results_prop_aps <- do.call("rbind", list(so2nox_comparison, noxr_comparison, 
                                          co2o2_comparison, h2o_comparison))
```

```{r }
# create table of results
kable(results_prop_aps, col.names = c("Test", "# RATAS", "# Pass 4QTRS", 
                                      "# Pass APS 4QTRS", "# Pass 2QTRS", "# Pass APS 2QTRS",
                                      "# Pass 4QTRS", "# Pass APS 4QTRS", "# Pass 2QTRS", 
                                      "# Pass APS 2QTRS", "# Passed", "% Passed", "# Change from Pass to Fail",
                                      "# Change from 4QTR to 2QTR", "% Change from 4QTR to 2QTR")) %>% 
  kable_styling(bootstrap_options = c("bordered","striped", "hover", "condensed", "responsive"), full_width=TRUE, position="left", font_size = 12) %>% 
  row_spec(row = 0, extra_css = "background-color: #2E8540; color: #fff;") %>%
  column_spec(column = 1, bold = TRUE) %>%
  column_spec(column = 2, extra_css = "border-right:2px solid darkgrey;") %>%
  column_spec(column = 7, extra_css = "border-left:2px solid darkgrey;") %>%
  column_spec(column = 13, extra_css = "border-left:2px solid darkgrey;") %>%
  add_header_above(c(" "=2, "Current Standards"=4, "Proposed Standards"=6, "Change"=3), extra_css = "background-color: #0071BC; color: #fff;") #%
```



